---
title: "The GLATOS package"
author: "Ryan Gosse"
date: "Dec 3, 2019"
output: html_document
---
## The GLATOS package

`glatos` is an R package with functions useful to members of the Great Lakes Acoustic Telemetry Observation System (http://glatos.glos.us). Functions may be generally useful for processing, analyzing, simulating, and visualizing acoustic telemetry data, but are not strictly limited to acoustic telemetry applications.

Follow and contribute to the development of the `glatos` package at its repository: https://gitlab.oceantrack.org/GreatLakes/glatos

```{r message=FALSE}
library(glatos)
library(tidyverse)

dets_file <- file.path('../data', 'detections.csv')
dets <- read_glatos_detections(dets_file)
dets <- false_detections(dets, tf = 3600)
```
```{r dets, echo=FALSE}
dets <- false_detections(dets, tf = 3600)
```
## Parse GLATOS results with dplyr

Now that we've got our filtered dataframe, let's bring back our Tidyverse tools and peek at what's in our output dataframe:

``` {r}
dets_filtered <- dets %>% filter(passed_filter != FALSE)
head(dets_filtered)
```

```{r}
# Just the ones that didn't pass the default time-based filter
dets_rejected <- dets %>% filter(passed_filter == FALSE)
head(dets_rejected)
```

## What do we see on manual review of the filtered-out detections?