---
title: "Additional VTrack"
author: "Ryan Gosse"
date: "Dec 3, 2019"
output: html_document
---
[`VTrack`](https://vinayudyawer.github.io/ATT/docs/ATT_Vignette.html "VTrack Reference") has some cool tools including COA and estimating activity spaces from passive telemetry data. This package has a suite of functions that help users follow a workflow to efficiently setup, analyse and visualise telemetry data. For more details on the workflow check out [this vignette](https://vinayudyawer.github.io/ATT/docs/ATT_Vignette.html).

# Setup data

The primary object `VTrack` uses is called an ATT object. This object houses the three core datasets required to analyse telemetry data:

1. Detections
2. Receiver deployment metadata
3. Tag metadata

Apart from these datasets, the ATT object also stores information on the Geographic Coordinate Reference System for that dataset. GLATOS is capable of exporting OTN data to a format readable by VTrack. Conversion of OTN data requires data from the [OTN ERDDAP server](https://members.oceantrack.org/erddap/tabledap/index.html?page=1&itemsPerPage=1000) Data can be found at the following links: [Animals](https://members.oceantrack.org/erddap/tabledap/otn_aat_animals.html), [Receivers](https://members.oceantrack.org/erddap/tabledap/otn_aat_receivers.html), and [Tag Releases](https://members.oceantrack.org/erddap/tabledap/otn_aat_tag_releases.html).

Lets load some data from the `glatos` package, remove false detections and set it up to use in `VTrack`


```{r message=FALSE}
library(tidyverse)
library(glatos)
library(plotly)
```

```{r}
wal_det_file <- system.file("extdata", "walleye_detections.csv", package = "glatos")
walleye_detections <- read_glatos_detections(wal_det_file)

rec_file <- system.file("extdata", "sample_receivers.csv",  package = "glatos")
rcv <- read_glatos_receivers(rec_file)

# find and remove false detections
filtered_detections <-
  walleye_detections %>% 
  false_detections(tf = 3600) %>% 
  filter(passed_filter != FALSE)

# Now we can convert the GLATOS filtered data into a ATT object to be used in VTrack
```

```{r walleye_detections} 
ATTdata <- glatos::convert_glatos_to_att(walleye_detections, rcv)

attr(ATTdata, "CRS") <- sp::CRS("+init=epsg:4326")

head(ATTdata)
```

```{r ATTdata}
library(VTrack)
aba <- abacusPlot(ATTdata, det.col = "red")

# Abacus plot by station
aba_facet <- abacusPlot(ATTdata, det.col = "red", facet = T)
```

```{r aba, aba_facet}
print(aba)
print(aba_facet)
```
```{r}
library(lubridate, verbose = FALSE, warn.conflicts = FALSE)

abacus_data <- 
  ATTdata$Tag.Detections %>% 
  left_join(ATTdata$Tag.Metadata,  by = "Transmitter") %>% 
  mutate(date = date(Date.Time)) %>% 
  group_by(Tag.ID, date, Sci.Name) %>% 
  summarise(Num.Det = n())

abacus_data %>% 
  ggplot(aes(x = date, y = factor(Tag.ID), size = Num.Det, color = Sci.Name)) +
  geom_point() +
  theme_linedraw() +
  labs(title = "Daily detection plot", x = "Date", y = "Tag ID", color = "Species", size = "Number of\ndetections")
```

```{r}
detSum <- detectionSummary(ATTdata, sub = "%Y-%W")

detSum

# Overall metrics
head(detSum$Overall)

# Weeky metrics
head(detSum$Subsetted)
```

```{r detSum}
detSum$Subset %>% 
  mutate(Date = as.Date(paste(subset, "01", sep = "-"), "%Y-%W-%w")) %>% 
  ggplot(aes(x = Date, y = Number.of.Detections, colour = factor(Tag.ID))) +
  geom_line() + geom_point() +
  labs(title = "Number of Detections over time", x = "Date", y = "Number of Detections", colour = "Tag ID") +
  theme_linedraw()
```


```{r}
dispSum <- dispersalSummary(ATTdata)

head(dispSum)

dispSum %>%
  mutate(Date = date(Date.Time),
         Tag.ID = factor(Tag.ID)) %>% 
  group_by(Date, Tag.ID) %>% 
  summarise(Max.Disp.km = max(Consecutive.Dispersal)) %>%
  ggplot(aes(x = Max.Disp.km, y=..scaled..,  fill = Tag.ID, color = Tag.ID)) +
  geom_density(alpha=0.1) +
  geom_rug(sides = "t", aes(y = Max.Disp.km, color = Tag.ID)) +
  ylim(0,1) +
  labs(x = "Maximum distance traveled per day (km)",
       y = "Relative frequency of daily movements",
       fill = "Tag ID", color = "Tag ID") +
  theme_linedraw()
```

```{r}
library(sf)
library(maptools)
library(mapview)

base <-
  ATTdata$Station.Information %>% 
  st_as_sf(coords=c("Station.Longitude", "Station.Latitude"), crs = 4326) %>% 
  mapview(alpha.regions = 0, color = "grey", cex = 3, legend = F, homebutton = F, layer.name = "Receiver Stations")


## Now lets use the COA() function to calculate hourly centers of activity
COAdata <- 
    COA(ATTdata, 
        timestep = 60) ## timestep used to estimate centers of activity (in minutes)




## Now lets add our centers of activity for each individual on the map
COAmap <- 
  base + 
  COAdata %>% 
  st_as_sf(coords = c("Longitude.coa", "Latitude.coa"), crs = 4326) %>% 
  mapview(zcol = "Tag.ID", burst = T, alpha = 0, cex = 3, legend = F)

COAmap
```

```{r COAdata, message=FALSE, warning=FALSE}
proj<-CRS("+init=epsg:3174")
mcp_est <- HRSummary(COAdata,
                     projCRS = proj,
                     type = "MCP",
                     cont = 100,
                     sub = "%Y-%W")

head(mcp_est)

## Lets have a look at how 100% MCP area changes over the tagging period
mcp_est$Subset %>% 
  mutate(Date = as.Date(paste(subset, "01", sep = "-"), "%Y-%W-%w")) %>% 
  ggplot(aes(x = Date, y = MCP.100/1e6, colour = factor(Tag.ID))) +
  geom_line() + geom_point() +
  labs(x = "Date", y = "Area of MCP (km2)", colour = "Tag ID") +
  theme_linedraw()
```

```{r message=FALSE}
BBkud_est <- HRSummary(COAdata,
                     projCRS = proj,
                     h = 200,
                     type = "BBKUD",
                     cont = c(20,50,95),
                     storepoly = TRUE,
                     sub = "%Y-%W")
                     
summary(BBkud_est)

# Overall Brownian bridge UD estimates for full tagging period
head(BBkud_est$Overall)

# Weekly Brownian bridge UD estimates for the three tagged walleye
head(BBkud_est$Subsetted)
```

```{r BBkud_est}
fullstack <-
  unlist(BBkud_est$Spatial.Objects)[grep("*_full", names(unlist(BBkud_est$Spatial.Objects)))]

names(fullstack) <-
  unlist(lapply(strsplit(names(fullstack), "[.]"), `[[`, 1))
```

```{r fullstack}
library(raster, verbose = FALSE, warn.conflicts = FALSE)

## Lets plot the overall BBKUD for Tag.ID `153`
fulltag <- fullstack$`153`
values(fulltag)[values(fulltag) > 95] <- NA

## Raster plot
plot(fulltag, zlim = c(0, 100))

## Mapview plot
base + 
  mapview(fulltag)
```

```{r}
library(leaflet, verbose = FALSE, warn.conflicts = FALSE)

## Full KUD for all tagged animals
fullmap <- COAmap@map

for (i in 1:length(fullstack)) {
  tempras<-disaggregate(fullstack[[i]], fact=3, method='bilinear')
  values(tempras)[values(tempras) >95] <-NA
  fullmap <- 
    fullmap %>% 
    addRasterImage(tempras, opacity = 0.7, group = names(fullstack)[i])
}

fullmap <- 
  fullmap %>%
  addLayersControl(
    baseGroups = names(fullstack),
    overlayGroups = "Receiver Stations",
    options = layersControlOptions(collapsed = FALSE)
  )
```